Kaluza--Klein Cosmology from five-dimensional Lovelock--Cartan Theory

### 2016-11-15 KKosmology.tex
Third (possible last) modification after referee report

\documentclass[aps,prd,12pt,superscriptaddress,showpacs,showkeys,longbibliography,reprint,nofootinbib]{revtex4-1}
\usepackage{calrsfs} 
\usepackage{amsmath,amsthm,latexsym,amssymb,amsfonts}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{float}
\usepackage[%
  colorlinks=true,
  urlcolor=blue,
  linkcolor=blue,
  citecolor=blue
]{hyperref}
\usepackage{etoolbox}
\usepackage{breqn}

@@

@@

We used the most general ansatz compatible with the cosmological symmetries (see Appendix~\ref{homotropic}) to find solutions of the Friedmann--Robertson--Walker class in the four-dimensional theory. The field equations for the cosmological ansatz ensure that the solutions with nontrivial torsion have vanishing $\alpha^{ab}$ and $\beta^a$. The solutions are parametrized by $u_\pm$, related to the fundamental couplings of the theory through the Eq.~\eqref{u}, and exhibit three different sectors depending on whether this parameter is positive, negative or zero. The behavior of the universes described by our solutions is four-fold: (i)~eternal expansion, (ii)~eternal contraction, (iii)~initial expansion with asymptotic contraction, and (iv)~bouncing. %We show the typical evolution of the scale factor for each case in Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}. 
In some cases, the solutions do not start from an initial singularity. The size of the extra dimension is driven by the scalar field $\phi(t)$ and its behavior at late times is compatible with a dynamical mechanism for compactification, at certain regions of the parameter space (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).

>>>>

\documentclass[aps,prd,12pt,superscriptaddress,showpacs,showkeys,reprint,nofootinbib]{revtex4-1}
\usepackage{calrsfs} 
\usepackage{amsmath,amsthm,latexsym,amssymb,amsfonts}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage{float}
\usepackage[%
  colorlinks=true,
  urlcolor=blue,
  linkcolor=blue,
  citecolor=blue
]{hyperref}
\usepackage{etoolbox}
\usepackage{breqn}
\allowdisplaybreaks

@@

Although there are several generalizations of gravitational models
which include higher order terms in curvature and torsion---in either
four or higher
dimensions~\cite{Puetzfeld:2004yg,Baekler:2011jt,Fabbri:2012qr}---, we
highlight that in our model the presence of higher order terms in
curvature and torsion are kept minimal, in the sense that we introduce
a sole extra coupling constant in our model (that accompanying the
Gauss--Bonnet term in five dimensions).
%% In the effective theory, we consider only the matter obtained from the
%% compactification process.
Moreover, the fact that our model is based in the Lovelock--Cartan
gravity assures that the field equations are first order differential
equations (since the Lagrangian can be written without the use of the
Hodge dual), which is not the case in most gravitational models
with higher order curvature and torsion.

@@

We used the most general ansatz compatible with the cosmological
symmetries (see Appendix~\ref{homotropic}) to find solutions of the
Friedmann--Robertson--Walker class in the four-dimensional
theory. Despite we restrict ourselves to the class of theories with
$\Delta > 0$, in Ref.~\cite{Canfora:2014iga} the case of negative
$\Delta$ is considered (without torsion). The field equations for the
cosmological ansatz ensure that the solutions with nontrivial torsion
have vanishing $\alpha^{ab}$ and $\beta^a$. The solutions are
parametrized by $u_\pm$, related to the fundamental couplings of the
theory through the Eq.~\eqref{u}, and exhibit three different sectors
depending on whether this parameter is positive, negative or zero. The
behavior of the universes described by our solutions is four-fold:
(i)~eternal expansion, (ii)~eternal contraction, (iii)~initial
expansion with asymptotic contraction, and (iv)~bouncing.
%We show the typical evolution of the scale factor for each case in
%Figs.~\ref{a_u_pos}, ~\ref{a_u_0} and~\ref{a_u_neg}.

Among the different cases, there are some solutions which do not start
from an initial singularity, however this is not the generic
behavior. Moreover, all of our cosmological solutions are free of
the future singularities categorized in Ref.~\cite{Shojai:2015dpa}. In
addition, the size of the extra dimension is driven by the scalar
field $\phi(t)$ and its behavior at late times is compatible with a
dynamical mechanism for compactification, at certain regions of the
parameter space (see Figs.~\ref{phi_u_0} and~\ref{phi_u_neg}).
